# Multilevel modeling in R

Yay, summer is over and we're back to doing methods :)
This file is not mandatory - but it will help you think about interpreting the outputs of multilevel models. Some of the questions may be difficult, so ask as many questions as you want to!


Start by loading in the data
```{r}
sim_data <- read.csv("sim_data.csv")
```

The (made up) data describes 100 participants' scores at a memory task (how many items they remembered correctly). Some have practiced it beforehand, indicated by the column "expertise", others are completely new to the task. Each participant has 3 trials to improve (so we have repeated measures!).


Lets do some models to look into the data (do try to understand the code)

```{r}
#first we fit a model with intercepts for each subject
memory_model <- lmer(score ~ expertise + trial_number + (1| ids), data = sim_data)

#and save the summary
memory_s <- summary(memory_model)

#what is in $coefficients? 
memory_s$coefficients 
#(Bonus question to remember what we did last semester: 
## are the estimates significant?) (don't get stuck if you don't remember)

#what does ranef() output?
subject_level <- memory_model %>% ranef() 

#look at the distribution of random intercepts - what shape is it?
subject_level$ids[,1] %>% hist()

#look at the standard deviation of the random intercepts
test_s$varcor
```
Here are some questions; you can ask Lydia if you are confused or puzzled

*Q1:* what does the group level intercept (13.2) represent in this model?
*A:* 

*Q2:* what does the subject level intercepts represent in this model? For instance, what does it mean that the modeled random intercept for the first participant is -0.62?
*A:*


*Cool fact:*
Because random effects are normally distributed, we can use their standard deviation to say what range most subjects' intercepts will be within (according to our model): the group level intercept plus or minus 2 standard deviations of the subject level intercepts. 

Compare this with the classical regression model, where we only model the group level intercept: in that case, we know that participants are different, but can't say anything about how different.


### Exercise 1
Your turn to model now! Maybe some participants improve faster than others, so edit the model from above to also include a random slope for trial_number for each participant. 
```{r}
#model goes here
```
Look at the summary(model)$varcor. Can you describe (or make a guess about) what each of the four numbers mean?
*A:*

Take a look at summary(model)$coefficients as well. Looking at both the group level slope for trial_number and the sd of subject level slopes, does the model predict that all people improve, or do some get negative slopes?
*A:*

### Exercise 2
Given the experimental design explained briefly at the start, would it make sense to add a random slope for expertise (instead of trial number) for each participant? What happens if you try? 

```{r}

```



Ending note: In a few weeks you will know why modeling this type of data ("counts") with a linear model is not ideal, and what might be a better choice instead - Then you can look back on this and feel super smart :)
